
---
subtitle: "TMA4268 Statistical Learning V2019"
title: "Compulsory exercise 3"
author: "Sara Elise Wøllo)"
date: "19 4 2020"
output: 
 # html_document
  pdf_document
---
  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3)
```

```{r rpackages,eval=TRUE,echo=FALSE}
# install.packages("knitr") #probably already installed
# install.packages("rmarkdown") #probably already installed
# install.packages("ggplot2") #plotting with ggplot
# install.packages("ggfortify")  
# install.packages("MASS")  
# install.packages("dplyr")  
library(knitr)
library(rmarkdown)
library(ggplot2)
library(ggfortify)
library(MASS)
library(dplyr)
```


# Problem 1


## a)

## b)

## c)

## d)



# Problem 2

```{r}
id <- "1CA1RPRYqU9oTIaHfSroitnWrI6WpUeBw"  # google file ID
d.corona <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", 
    id), header = T)
```

```{r}
count(d.corona, country, deceased)
count(d.corona, sex, deceased)
count(d.corona, country, sex, deceased)
```
We see that there are 14 (5 female (f), 9 male (m)) deceased in France, 2 (1 (f), 1 (m)) in Indonesia, 3 (0 (f), 3 (m)) in Japan and 26 ( 8(f), 18 (m)) in Korea. 
There are 14 females deceased and 31 males deceased. 



## a)


## b) 


## c)

## d)

## e)

## f)




# Problem 3

## a)

## b)

## c)

## d)

## e)




# Problem 4

## a)

## b)

## c)



# Problem 5

## a)
 i) True, ii) True, iii) False, iv) Unsure (You can use reg, but is it a method for reg?)

## b)
i) False ii) True iii)  False iv) True?

## c) Single choice

True: iv) ? [it increases, but can we say that it is a steady increase?]

## d) Single choice

True: ii) [is it always true? In general, it performs worse with more p]

## e) Single choice - ???
i)  ii)  iii)   iv) 

## f)
i) True ii) False ? (can happen, but not always true. as sensitivity increases, specificity doesn't HAVE to go down)  iii)  False iv) True

## g)
i) True  ii) False iii) True  iv) True
